Wigglyhedra

### 2024-02-27 draft/wigglyhedra.tex
improve Cambrian

\usepackage[T1]{fontenc}
\usepackage{enumerate, amsmath, amsfonts, amssymb, amsthm, mathrsfs, wasysym, graphics, graphicx, xcolor, url, hyperref, hypcap, shuffle, xargs, multicol, overpic, pdflscape, multirow, hvfloat, minibox, accents, array, multido, xifthen, a4wide, ae, aecompl, blkarray, pifont, mathtools, etoolbox, dsfont}
\usepackage{marginnote}
\hypersetup{colorlinks=true, citecolor=darkblue, linkcolor=darkblue}
\usepackage[all]{xy}
\usepackage[bottom]{footmisc}
\usepackage{tikz}
%\usepackage{tkz-graph}
%\usepackage{tikz-qtree}
\usetikzlibrary{trees, decorations, decorations.markings, shapes, arrows, matrix, calc, fit, intersections, patterns, angles}
\usepackage[external]{forest}
%\tikzexternalize
\graphicspath{{figures/}{figures/nodes/}}
\makeatletter\def\input@path{{figures/}}\makeatother
\usepackage{caption}
\captionsetup{width=\textwidth}
\renewcommand{\topfraction}{1} % possibility to have one page of pictures
\renewcommand{\bottomfraction}{1} % possibility to have one page of pictures
\usepackage[noabbrev,capitalise]{cleveref}
\usepackage[export]{adjustbox}
\usepackage{ulem}\normalem

@@

The following definition is illustrated in \cref{fig:incompatible2}.
%
\begin{figure}
\centering
\includegraphics[scale=1.3]{incompatible2}
\caption{The arcs~$\alpha$ and~$\alpha'$ of \cref{def:wigglyIncreasingFlipGraph}.}
\label{fig:incompatible2}
\end{figure}

@@

Finally, we will need the following observation, illustrated in \cref{fig:incompatible3}.
%
\begin{figure}
\centering
\includegraphics[scale=1.3]{incompatible3}
\caption{The arcs~$\beta$ and~$\beta'$ of \cref{prop:uerp}.}
\label{fig:incompatible3}
\end{figure}

@@

%The following statements are illustrated on \cref{fig:wigglyLattice}\,(right) when~${n = 2}$.
%
\begin{figure}
\centering
\includegraphics[scale=1.1]{wigglyFlipGraph}
\caption{The wiggly lattice~$\wigglyLattice_2$ on wiggly pseudotriangulations (left) and on wiggly permutations (right).}
\label{fig:wigglyLattice}
\end{figure}

@@

We now construct the wigglyhedron illustrated in \cref{fig:wigglyhedron} when~${n = 2}$.
%
\begin{figure}[t]
\centering
\input{wigglyhedron}
\caption{The wigglyhedron~$\wigglyhedron_2$.}
\label{fig:wigglyhedron}
\end{figure}

@@

In this section, we fix a signature~$\delta \in \{+,-\}^n$. By convention, we set~$\delta_0 = \delta_{n+1} = 0$.

\begin{definition}
The \defn{$\delta$-gon} is a convex $(n+2)$-gon with a point with abscissa~$i$ and with ordinate of the same sign as~$\delta_i$  for all~$0 \le i \le n+1$.
A \defn{$\delta$-triangulation} is a triangulation of the $\delta$-gon.
The \defn{$\delta$-triangulation lattice} is the transitive closure of the slope increasing flip graph on $\delta$-triangulations.
%If~$T$ and~$T'$ are two $\delta$-triangulations with~$T \ssm \{d\} = T' \ssm \{d'\}$, then the flip from~$T$ to~$T'$ is \defn{increasing} if the slope of~$d$ is smaller than the slope of~$d'$.
\end{definition}

\begin{definition}
A \defn{$\delta$-permutation} is a permutation of~$[n]$ which avoids~$\cdots ik \cdots j \cdots$ with~$i < j < k$ and~$\delta_j = -$, and~$\cdots j \cdots ik \cdots$ with~$i < j < k$ and~$\delta_j = +$.
The \defn{$\delta$-permutation lattice} is the sublattice of the weak order induced by $\delta$-permutations.
\end{definition}

@@

@@

\begin{definition}
Fix a point set~$P$ of the plane, and an arbitrary total order~$<$ on~$P$.
A \defn{wiggly arc} is a quadruple~$(p,q,R,S)$ where~$p < q \in P$ and the sets~$R$ and~$S$ form a partition of the points of~$P$ located in the open segment joining~$p$ to~$q$.
Two wiggly arcs~$(p,q,R,S)$ and~$(p',q',R',S')$ are \defn{crossing} if 
\begin{itemize}
\item either the segments~$[p,q]$ and~$[p',q']$ cross, 
\item or~$(R \cap S') \cup (\{p,q\} \cap S') \cup (R \cap \{p',q'\}) \ne \varnothing \ne (R' \cap S) \cup (\{p',q'\} \cap S) \cup (R' \cap \{p,q\})$,
\end{itemize}
A set~$X$ of wiggly arcs is \defn{pointed} if for any~$p \in P$, the wiggly arcs of~$X$ with an endpoint at~$p$ generate a pointed cone.
The \defn{wiggly complex}~$\wigglyComplex_P$ is the simplicial complex of pairwise pointed and non-crossing subsets of wiggly arcs.
Note that the boundary wiggly arcs are irrelevant, which allows us to consider a reduced wiggly complex~$\wigglyComplex_P$ induced by internal wiggly arcs.
A \defn{wiggly pseudotriangulation} of~$P$ is a facet of~$\wigglyComplex_P$.
The \defn{wiggly flip graph}~$\wigglyFlipGraph_P$ is the adjacency graph of the facets of~$\wigglyComplex_P$.
Note that, by construction, $\wigglyFlipGraph_P$ is regular.
See \cref{fig:wigglyComplexSquarre} for an illustration.
%
\begin{figure}[!h]
\centering
\includegraphics[scale=1]{wigglyComplexSquare}
\caption{The wiggly complex~$\wigglyComplex_P$ (left) and the wiggly flip graph~$\wigglyFlipGraph_P$ (right) of a point set~$P$.}
\label{fig:wigglyComplexSquarre}
\end{figure}
\end{definition}

>>>>

\usepackage[T1]{fontenc}
\usepackage{enumerate, amsmath, amsfonts, amssymb, amsthm, mathrsfs, wasysym, graphics, graphicx, xcolor, url, hyperref, hypcap, shuffle, xargs, multicol, overpic, pdflscape, multirow, hvfloat, minibox, accents, array, multido, xifthen, a4wide, ae, aecompl, blkarray, pifont, mathtools, etoolbox, dsfont}
\usepackage{marginnote}
\hypersetup{colorlinks=true, citecolor=darkblue, linkcolor=darkblue}
\usepackage[all]{xy}
\usepackage[bottom]{footmisc}
\usepackage{tikz}
%\usepackage{tkz-graph}
%\usepackage{tikz-qtree}
\usetikzlibrary{trees, decorations, decorations.markings, shapes, arrows, matrix, calc, fit, intersections, patterns, angles, cd}
\usepackage[external]{forest}
%\tikzexternalize
\graphicspath{{figures/}{figures/nodes/}}
\makeatletter\def\input@path{{figures/}}\makeatother
\usepackage{caption}
\captionsetup{width=\textwidth}
\renewcommand{\topfraction}{1} % possibility to have one page of pictures
\renewcommand{\bottomfraction}{1} % possibility to have one page of pictures
\usepackage[noabbrev,capitalise]{cleveref}
\usepackage[export]{adjustbox}
\usepackage{ulem}\normalem

@@

The following definition is illustrated in \cref{fig:incompatible2}.
%
\begin{figure}
\centerline{\includegraphics[scale=1.3]{incompatible2}}
\caption{The arcs~$\alpha$ and~$\alpha'$ of \cref{def:wigglyIncreasingFlipGraph}.}
\label{fig:incompatible2}
\end{figure}

@@

Finally, we will need the following observation, illustrated in \cref{fig:incompatible3}.
%
\begin{figure}
\centerline{\includegraphics[scale=1.3]{incompatible3}}
\caption{The arcs~$\beta$ and~$\beta'$ of \cref{prop:uerp}.}
\label{fig:incompatible3}
\end{figure}

@@

%The following statements are illustrated on \cref{fig:wigglyLattice}\,(right) when~${n = 2}$.
%
\begin{figure}
\centerline{\includegraphics[scale=1.1]{wigglyFlipGraph}}
\caption{The wiggly lattice~$\wigglyLattice_2$ on wiggly pseudotriangulations (left) and on wiggly permutations (right).}
\label{fig:wigglyLattice}
\end{figure}

@@

We now construct the wigglyhedron illustrated in \cref{fig:wigglyhedron} when~${n = 2}$.
%
\begin{figure}[t]
\centerline{\input{wigglyhedron}}
\caption{The wigglyhedron~$\wigglyhedron_2$.}
\label{fig:wigglyhedron}
\end{figure}

@@

In this section, we fix a signature~$\delta \in \{+,-\}^n$. By convention, we set~$\delta_0 = \delta_{n+1} = 0$.
We connect the $\delta$-Cambrian lattice~\cite{Reading-CambrianLattices} to the wiggly lattice of \cref{sec:combinatorics} and the $\delta$-associahedron of~\cite{HohlwegLange} to the wigglyhedron of \cref{sec:geometry}.
\cref{def:CambrianTriangulations,def:CambrianPermutations,def:CambrianWigglyPseudotriangulations,def:CambrianWigglyPermutations,prop:CambrianLattice} are illustrated in \cref{fig:wigglyCambrian}.

\begin{figure}
\centerline{\includegraphics[scale=1.1]{wigglyCambrian1}}
\vspace{-.3cm}
\centerline{\includegraphics[scale=1.1]{wigglyCambrian2}}
\vspace{-.5cm}
\centerline{\includegraphics[scale=1.1]{wigglyCambrian3}}
\caption{Cambrian lattices as intervals of wiggly lattices.}
\label{fig:wigglyCambrian}
\end{figure}

\begin{definition}
\label{def:CambrianTriangulations}
The \defn{$\delta$-gon} is a convex $(n+2)$-gon with a point with abscissa~$i$ and with ordinate of the same sign as~$\delta_i$  for all~$0 \le i \le n+1$.
A \defn{$\delta$-triangulation} is a triangulation of the $\delta$-gon.
The \defn{$\delta$-triangulation lattice} is the transitive closure of the slope increasing flip graph on $\delta$-triangulations.
%If~$T$ and~$T'$ are two $\delta$-triangulations with~$T \ssm \{d\} = T' \ssm \{d'\}$, then the flip from~$T$ to~$T'$ is \defn{increasing} if the slope of~$d$ is smaller than the slope of~$d'$.
\end{definition}

\begin{definition}
\label{def:CambrianPermutations}
A \defn{$\delta$-permutation} is a permutation of~$[n]$ which avoids~$\cdots ik \cdots j \cdots$ with~$i < j < k$ and~$\delta_j = -$, and~$\cdots j \cdots ki \cdots$ with~$i < j < k$ and~$\delta_j = +$.
The \defn{$\delta$-permutation lattice} is the sublattice of the weak order induced by $\delta$-permutations.
\end{definition}

@@

\begin{proof}
We only describe the following bijections
\begin{center}
\begin{tikzcd}[column sep=3cm]
\text{$\delta$-triangulations} \arrow[d] & \text{$\delta$-permutations} \arrow[l] \\
\text{$\delta$-wiggly pseudotriangulations} \arrow[r] & \text{$\delta$-wiggly permutations} \arrow[u]
%\text{$\delta$-triangulations} \arrow[d, "\lambda"'] & \text{$\delta$-permutations} \arrow[l, "\zeta"'] \\
%\text{$\delta$-wiggly pseudotriangulations} \arrow[r, "\mu"'] & \text{$\delta$-wiggly permutations} \arrow[u, "\nu"']
\end{tikzcd}
\end{center}
illustrated in \cref{fig:wigglyCambrian}.
We let the readers convince themselves that they are well-defined, that they compose to the identity, and that they preserve the flip structure. (We also invite the reader to work out the explicit description of the 8 missing arrows.)
For~$I \subseteq [n]$, we set~$I^+ \eqdef \set{i \in I}{\delta_i = {+}}$ and~$I^- \eqdef \set{i \in I}{\delta_i = {-}}$.

\para{From $\delta$-triangulations to $\delta$-wiggly pseudotriangulations}
Replace each diagonal~$(i,j)$ of the $\delta$-gon by the wiggly arc~$(0, j, [i] \cup {]i,j[}^-, {]i,j[}^+)$ if~$\delta_i \ge 0$ and~$(0, j, {]i,j[}^-, [i] \cup {]i,j[}^+)$ if~$\delta_i \le 0$.

\para{From $\delta$-wiggly pseudotriangulations to $\delta$-wiggly permutations}
Already described in \cref{def:bijection1}.

\para{From $\delta$-wiggly permutations to $\delta$-permutations}
For all~$j \in [n]$, erase~$2j-1$ and replace~$2j$ by~$j$ if~$\delta_j = {+}$, and erase~$2j$ and replace~$2j-1$ by~$j$ if~$\delta_j = {-}$.

\para{From $\delta$-permutations to $\delta$-triangulations}
Already described in~\cite{Reading-CambrianLattices}.
Given a permutation~$\pi$, construct the union of the upper hulls of the point sets~$[n]^- \symdif \pi([j])$ for all~$j \in \{0, \dots, n\}$.
\end{proof}

@@

\begin{definition}
Fix a point set~$P$ of the plane, and an arbitrary total order~$<$ on~$P$.
A \defn{wiggly arc} is a quadruple~$(p,q,R,S)$ where~$p < q \in P$ and the sets~$R$ and~$S$ form a partition of the points of~$P$ located in the open segment joining~$p$ to~$q$.
Two wiggly arcs~$(p,q,R,S)$ and~$(p',q',R',S')$ are \defn{crossing} if 
\begin{itemize}
\item either the segments~$[p,q]$ and~$[p',q']$ cross, 
\item or~$(R \cap S') \cup (\{p,q\} \cap S') \cup (R \cap \{p',q'\}) \ne \varnothing \ne (R' \cap S) \cup (\{p',q'\} \cap S) \cup (R' \cap \{p,q\})$,
\end{itemize}
A set~$X$ of wiggly arcs is \defn{pointed} if for any~$p \in P$, the wiggly arcs of~$X$ with an endpoint at~$p$ generate a pointed cone.
The \defn{wiggly complex}~$\wigglyComplex_P$ is the simplicial complex of pairwise pointed and non-crossing subsets of wiggly arcs.
Note that the boundary wiggly arcs are irrelevant, which allows us to consider a reduced wiggly complex~$\wigglyComplex_P$ induced by internal wiggly arcs.
A \defn{wiggly pseudotriangulation} of~$P$ is a facet of~$\wigglyComplex_P$.
The \defn{wiggly flip graph}~$\wigglyFlipGraph_P$ is the adjacency graph of the facets of~$\wigglyComplex_P$.
Note that, by construction, $\wigglyFlipGraph_P$ is regular.
See \cref{fig:wigglyComplexSquarre} for an illustration.
%
\begin{figure}[!h]
\centerline{\includegraphics[scale=1]{wigglyComplexSquare}}
\caption{The wiggly complex~$\wigglyComplex_P$ (left) and the wiggly flip graph~$\wigglyFlipGraph_P$ (right) of a point set~$P$.}
\label{fig:wigglyComplexSquarre}
\end{figure}
\end{definition}
